### 2017-05-16 paper/jfp-submission/evaluation.tex

\paragraph{4. Witnesses Assign Blame}
Finally, we present a simple heuristic that allows us to use witnesses
to \emph{automatically} assign blame for type errors.
%
We treat the values inside the stuck term as \emph{sources} of typing
constraints and the stuck term itself as a \emph{sink}, producing
a slice of the program that likely contains the error.
%
Using this heuristic, \toolname's witness are competitive with the
state-of-the-art localization tools \mycroft and \sherrloc.

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs
from a cohort of 46 students..
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
polymorphic and algebraic data types. % and expression evaluators.

>>>>

\paragraph{4. Witnesses Assign Blame}
Finally, we present a simple heuristic that allows us to use witnesses
to \emph{automatically} assign blame for type errors.
%
We treat the values inside the stuck term as \emph{sources} of typing
constraints and the stuck term itself as a \emph{sink}, producing
a slice of the program that likely contains the error.
%
Using this heuristic, \toolname's witnesses are competitive with the
state-of-the-art localization tools \mycroft and \sherrloc.

\subsection{Methodology}
\label{sec:methodology}
We answer the first two questions on two sets of ill-typed programs,
\ie\ programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages (CSE 130) course at UC San Diego.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB
% \# hidden for blind review),
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml\ programs
from a cohort of 46 students..
%
The second dataset --- widely used in the literature --- comes from a
graduate-level course at the University of Washington~\cite{Lerner2006-pj},
from which we extracted 284 ill-typed programs.
%
Both datasets contain relatively small programs, the largest being 348
SLoC; however, they demonstrate a variety of functional programming
idioms including (tail) recursive functions, higher-order functions,
and polymorphic and algebraic data types. % and expression evaluators.

### 2017-05-16 paper/jfp-submission/haskellListings.tex

\lstnewenvironment{ecode}
{\lstset{language=HaskellUlisses
				,numbers=left
                                ,frame=leftline
                                ,xleftmargin=7.5mm
				,moredelim=[is][\textbf]{==}{==}
				,moredelim=[is][\underbar]{__}{__}
				}}
{}

>>>>

\lstnewenvironment{ecode}
{\lstset{language=HaskellUlisses
        ,numbers=left
        ,frame=leftline
        ,xleftmargin=7.5mm
        ,escapeinside={(*@}{@*)}
}}
{}

### 2017-05-16 paper/jfp-submission/locating.tex

As discussed in \S~\ref{sec:methodology}, we recorded
each interaction of our students with the \ocaml top-level system.
%
In addition to collecting ill-typed programs, we collected
subsequent, fixed versions of the programs.
%
For each ill-typed program in a student's interaction trace, we identify
the student's fix by searching for the first type-correct program
that follows it in the trace.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

>>>>

As discussed in \S~\ref{sec:methodology}, we recorded
each interaction of our students with the \ocaml top-level system.
%
This means that, in addition to collecting ill-typed programs, we
collected subsequent, fixed versions of the same programs.
%
% For each ill-typed program in a student's interaction trace, we identify
% the student's fix by searching for the first type-correct program
% that follows it in the trace.
For each ill-typed program compiled by a student, we identify the student's
\emph{fix} by searching for the first type-correct program that the student
subsequently compiled.
%
We then use an expression-level \emph{diff}~\cite{Lempsink2009-xf} to
determine which sub-expressions changed between the ill-typed program
and the student's fix, and treat those expressions as the source of the
type error.

### 2017-05-16 paper/jfp-submission/overview.tex

\paragraph{Reduction Graph}
Our trace takes the form of a reduction graph, which records small-step
reductions in the context in which they occur.
%
% These graphs have two types of edges:
% %
% (1) ``steps-to'' edges that capture the small-step transition between
% two terms, and
% %
% (2) ``sub-term'' edges that capture the containment relation between two
% terms.
%
For example, evaluating the expression @1+2+3@ would produce the
graph in Figure~\ref{fig:simple-reduction-hi}.
%
\begin{figure}[t]
  \centering
  \includegraphics[height=2in]{simple.png}
  \caption{The reduction graph for \texttt{1+2+3}. The two edges
    produced by the transition from \texttt{1+2+3} to \hbox{\texttt{3+3}}
    are highlighted.}
\label{fig:simple-reduction-hi}
\end{figure}
%
Notice that when we transition from @1+2+3@ to @3+3@ we collect
both that edge \emph{and} an edge from the sub-term @1+2@ to @3@.
%
These additional edges allow us to implement two common debugging
operations \emph{post-hoc}: ``step into'' to zoom in on a specific
function call, and ``step over'' to skip over an uninteresting
sub-computation.

>>>>

\paragraph{Reduction Graph}
Our trace takes the form of a reduction graph, which records small-step
reductions in the context in which they occur.
%
% These graphs have two types of edges:
% %
% (1) ``steps-to'' edges that capture the small-step transition between
% two terms, and
% %
% (2) ``sub-term'' edges that capture the containment relation between two
% terms.
%
For example, evaluating the expression @1+2+3@ would produce the
graph in Figure~\ref{fig:simple-reduction-hi}.
%
\begin{figure}[t]
  \centering
  \includegraphics[height=2in]{simple.png}
  \caption{The reduction graph for \texttt{1+2+3}, highlighting the edges
    produced by reducing \texttt{1+2+3} to \hbox{\texttt{3+3}}.}
\label{fig:simple-reduction-hi}
\end{figure}
%
Notice that when we transition from @1+2+3@ to @3+3@ we collect
both that edge \emph{and} an edge from the sub-term @1+2@ to @3@.
%
These additional edges allow us to implement two common debugging
operations \emph{post-hoc}: ``step into'' to zoom in on a specific
function call, and ``step over'' to skip over uninteresting
computations.
